Emissions and Perceptions of Climate Change

### 2022-12-14 Code/Drafts/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Finished Final CAP DAP.

```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
```

```{r}
t.test(rate~country, data=emissions_rate, paired=TRUE)
#complete.cases means just use the ones without NAs
```

```{r}
t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
```

We can also compare the moving averages and compare the rates, even forecasting to future dates (like what will emissions be in 2050 or 2100 based on the current time series). This comparision will be visual for the moving averages and just thoughtful for comparing the forecast numbers and will not involve a specific statistical test.

#### Running the Test

First determine the minimum date
```{r}
min_date=min(emissions$year)
max_date=max(emissions$year)
```
I'm not sure you'll need the above code chunk

Turning our data into a time series

```{r}
emissions_ts<-ts(emissions$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
```

Lets do time series for the indiviual countries
```{r}
emissions_Denmark<- emissions %>% filter(country=="Denmark")
emissions_US<-emissions %>% filter(country=="United States")
emissions_Denmark_ts<-ts(emissions_Denmark$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
emissions_US_ts<-ts(emissions_US$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
```

Viewing the time series of each individual country to ensure its correct.
```{r, results='hide'}
both_ts<-cbind(emissions_Denmark_ts, emissions_US_ts)
head(both_ts) # ensuring it worked
plot(both_ts, xlab="year", ylab="emissions")
```

If you wish to plot time series using ggplot, you can use autoplot(timeseriesname) and save it as an object. The result will be a ggplot object that you can manipulate. Example:

```{r}
demo_plot<-autoplot(emissions_US_ts)
demo_plot+
  labs(x = "Year", y = "Greenhouse gas emissions")+
  theme_bw()
```

And if you want to plot two time series on the same graph using ggplot, you use cbind to bind your two time series objects into the same new data frame and then autoplot that data frame:

```{r}
both<-cbind(emissions_Denmark_ts, emissions_US_ts)
autoplot(both)
```

Here is our data smoothed over so we can see the trends without added noise. The bottom one I presume is denmark. We can see that it is decreasing more.

Making a moving average:
```{r}
Ma_emissions<- forecast::ma(emissions_ts, order=3, centre = TRUE)
Ma_emissions_Denmark<- forecast:: ma(emissions_Denmark_ts, order=3, centre=TRUE)
Ma_emissions_US<- forecast::ma(emissions_US_ts, order=3, centre=TRUE)

@@

Plotting both the moving average and the time series.

```{r, collapse=TRUE}
plot(emissions_ts)
lines(Ma_emissions, col="maroon", lwd=3)

plot(emissions_Denmark_ts)
lines(Ma_emissions_Denmark, col="orange", lwd=3)

plot(emissions_US_ts)
lines(Ma_emissions_US, col="blue")
```
Plotting both with autoplot/ggplot:

```{r}
USA<-cbind(emissions_US_ts, Ma_emissions_US)
autoplot(USA)
```

The red line is the data but smoothed, so we can see the overall trend of the emissions data combined. the other two are them separated, notice how similar Denmark's is to the combined data. Interesting.

***PUT RUNNING THE TIME SERIES HERE***
This will really be the paired t-test

@@

we will run another generalized linear model.

@@

## Survey data Impacting Emissions

@@

>>>>

```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
```
There is alot of noise, and no discernable trends, so lets look at it smoothed over:

```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
  geom_smooth() + theme_bw() #colour colors by country, and group tells R to 
#group the points by country so geom_line will draw lines connecting the 
#data by country.

```
Denmark's rate is more variable over time, but it stays below 0, after 1995, wherease the US is not consistently below zero until 2000. It also goes above zero at 1990. Therefore, Denmark started decreasing its emissions five years prior to the U.S. and the U.S.'s emissions might be increasing again in 2019. Given Denmark has a lower rate from 2010-2019, it may have a significantly lower rate than the U.S. Lets run a statistical test to verify.

```{r}
t.test(rate~country, data=emissions_rate, paired=TRUE)
#complete.cases means just use the ones without NAs
```
Looking at the entire range from 1990-2019, there is no significant difference in rate between Denmark and the U.S. Lets look at 2010-2019 given the differences in the funcitons listed above.

```{r, echo=FALSE}
emissions_rate_2010<- emissions_rate %>% 
  filter(emissions_rate$year>="2010-01-01")

@@

```{r, results='hold', echo=FALSE}
t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
#complete.cases means just use the ones without NAs
```

The p value is smaller at 0.083, however it is still not signifciant. So, we must accept the Null hypothesis. There is no significant difference in rate between Denmark and the United States.

#### Conclusion for Section One

The United States emits significantly more greenhouse gases per capita than Denmark, but there was not significant difference in how fast both countries decreased their emissions.

Now lets look at some examples that may explain why Denmark emits less greenhouse gases per capita.

@@

Making a Diverging Bar Graph
```{r}
fuels_diverging <- fuels %>% mutate(percent=if_else(response %in% c("Much less", "Somewhat less"), Number, -Number)) 
#everything in there 
```

```{r}
fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(percent, digits=1), "%"))
ggplot(fuels_diverging, aes(x=Country, y=Number, fill=response)) + 
  geom_col()+ #also makes a bar graph
  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
  coord_flip()+
  scale_x_discrete()+
  scale_fill_viridis_d()+
  labs(title="How much Fossil Fuels Should We Use", x=NULL, fill=NULL)+
 theme_minimal()+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
panel.grid=element_blank(),
legend.position="top")
```
```{r, echo=FALSE}
dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
fuels$num_people<- dumb_way #put that vector into here

fuels<- fuels %>% group_by(Country, response) %>% 
  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage

fuels$total<- round(fuels$total, digits=0) #round to whole number

fuels_numPeople<- fuels %>% select(response, Country, total)
```

```{r}

fuels_numPeople<-fuels_numPeople %>% filter(response != "Don’t know") #excluding don't know

fuels_numPeople$grouped_answer<- NA
more<- which( fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less" )
less<- which( fuels_numPeople$response== "Much more" |fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Refused" |fuels_numPeople$response== "Same amount as today")

  fuels_numPeople$grouped_answer[more]<- "more"
  fuels_numPeople$grouped_answer[less]<- "less"

fuels_numPeople<- fuels_numPeople %>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))

@@

```{r}
fuels.mat<-xtabs(grouped_number~Country, data=fuels_numPeople)
fuels.mat
chisq.test(fuels.mat)
```
The p value is 2.2e^-16, so we can reject the Null hypothesis. significantly more Americans think we should keep releasing the same amount of fossil fuels or more fossil fuels than Danes.

@@

## DIVERGING BAR GRAPH SAVING THIS

```{r}
beliefs_diverging <- beliefs %>% mutate(percent=if_else(response %in% c("Caused mostly by human activities"), Number, -Number)) 
#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
```

```{r}
# This will make our labels are positive numbers.

beliefs_diverging<- beliefs_diverging %>% mutate(percent=paste0(round(percent, digits=1), "%"))
ggplot(beliefs_diverging, aes(x=Country, y=Number, fill=response)) + 
  geom_col()+ #also makes a bar graph
  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
  coord_flip()+
  scale_x_discrete()+
  scale_fill_viridis_d()+
  labs(title="What do you Think Caused Climate Change", x=NULL, fill=NULL)+
 theme_minimal()+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
panel.grid=element_blank(),
legend.position="top")
